Chronic conditions and increased antibiotic consumption

### 2017-03-28 chronic-conditions/chronic-condition-drivers.Rmd
added mcc

```{r load_data}
all_denom = read_tsv('../bene_2011.tsv') %>%
  filter(age >= 65) %>%
  mutate(is_female=sex=='female') %>%
  select(bene_id, age, is_female)
chronic_conditions = read_tsv('cc_2011.tsv') %>%
  rename(bene_id=BENE_ID) %>%
  select(-STRCTFLG) %>%
  mutate_at(vars(AMI:HYPOTH), function(x) x==3)
usage = read_tsv('../pde_2011.tsv') %>%
  count(bene_id) %>% rename(n_claims=n)

na_to_0 = function(x) if_else(is.na(x), 0L, x)

usage_summary = denom %>%
  left_join(chronic_conditions, by='bene_id') %>%
  left_join(usage, by='bene_id') %>%
  mutate(n_claims=na_to_0(n_claims))

@@

res = lapply(conditions, function(cond) summarize_condition(usage_summary, cond)) %>%
  bind_rows() %>%
  left_join(condition_names, by='condition') %>%
  select(-condition) %>%
  rename(condition=condition_name)
```

@@

multivariate_res = usage_summary %>%
  select(n_claims, age, is_female, AMI:HYPOTH) %>%
  glm(multivariate_formula, data=., family='poisson') %>%
  tidy
```

@@

Thankfully, Alzheimer's correlates strongly with Alzheimer's and dementia, and prostatic hyperplasia and prostate cancer correlate strongly negatively with breast cancer and endometrial cancer. Unfortunately, the negative correlations make everything else harder to see, so here's the same thing but without the male conditions.

```{r fig.width=7, fig.height=7, cache=FALSE}
(!(rownames(tetra$rho) %in% c('CNCRPRST', 'HYPERP'))) %>%
  { tetra$rho[., .] } %>%
  heatmap.2(trace='none')
```

@@

@@

```{r copd}
usage_summary %>%
  group_by(COPD) %>%
  do(summary(.$n_claims) %>% tidy) %>%
  kable
```

>>>>

```{r load_data}
bene = read_tsv('../bene_2011.tsv') %>%
  filter(age >= 65) %>%
  mutate(is_female=sex=='female') %>%
  select(bene_id, age, is_female)
chronic_conditions = read_tsv('cc_2011.tsv') %>%
  rename(bene_id=BENE_ID) %>%
  select(-STRCTFLG) %>%
  mutate_at(vars(AMI:HYPOTH), function(x) x==3)
pde = read_tsv('../pde_2011.tsv') %>%
  count(bene_id) %>% rename(n_claims=n)

usage = bene %>%
  left_join(chronic_conditions, by='bene_id') %>%
  left_join(pde, by='bene_id') %>%
  replace_na(list(n_claims=0))

@@

res = lapply(conditions, function(cond) summarize_condition(usage, cond)) %>%
  bind_rows() %>%
  left_join(condition_names, by='condition') %>%
  select(-condition) %>%
  rename(condition=condition_name)
```

@@

multivariate_res = usage %>%
  select(n_claims, age, is_female, AMI:HYPOTH) %>%
  glm(multivariate_formula, data=., family='poisson') %>%
  tidy
```

@@

Thankfully, Alzheimer's correlates strongly with Alzheimer's and dementia, and prostatic hyperplasia and prostate cancer correlate strongly negatively with breast cancer and endometrial cancer. Unfortunately, the negative correlations make everything else harder to see, so here's the same thing but without the male conditions. I also leave out Alzheimer's and dementia, because it's extremely well-correlated with Alzheimer's and leads to some confusing setup.

```{r fig.width=7, fig.height=7, cache=FALSE}
(!(rownames(tetra$rho) %in% c('CNCRPRST', 'HYPERP', 'ALZHDMTA'))) %>%
  { tetra$rho[., .] } %>%
  heatmap.2(trace='none')
```

@@

### MCC correlations

```{r mcc}
mcc = function(true, pred) {
  ut = true - mean(true)
  up = pred - mean(pred)

  co = mean(ut * up)
  vt = mean(ut ** 2)
  vp = mean(up ** 2)

  if (vt * vp == 0) {
    NA
  } else {
    co / sqrt(vt * vp)
  }
}

mcc_cor = function(x) {
  x = as.matrix(x)
  labels = colnames(x)
  n = length(labels)
  out = matrix(nrow=n, ncol=n, dimnames=list(labels, labels))

  for (i in 1:n) out[i, i] = 1.0

  for (i in 1:(n-1)) {
    ui = x[,i] - mean(x[,i])
    vi = mean(ui ** 2)
    for (j in (i+1):n) {
      uj = x[,j] - mean(x[,j])
      vj = mean(uj ** 2)
      co = mean(ui * uj)
      val = co / sqrt(vi * vj)
      out[i, j] = val
      out[j, i] = val
    }
  }
  out
}

mccs = chronic_conditions %>%
  select(-bene_id) %>%
  mcc_cor
```

Here is a heatmap using the square roots of the Matthews correlation coefficients. (I used square roots because it emphasizes that there are two clusters. Using just raw values is not as nice to look at.) Interestingly, all the correlations are positive.

```{r fig.width=7, fig.height=7, cache=FALSE}
(!(rownames(mccs) %in% c('CNCRPRST', 'HYPERP', 'ALZHDMTA'))) %>%
  { mccs[., .] } %>%
  sqrt %>%
  gplots::heatmap.2(trace='none', symm=TRUE)
```

@@

```{r copd}
usage %>%
  group_by(COPD) %>%
  do(summary(.$n_claims) %>% tidy) %>%
  kable
```
